---
title: "DAATA VISUALIZATION"
author: "OLUWAFIKEMI"
date: "5/25/2022"
output: html_document
---

SCATTER PLOT
Customization of scatter plots is very much acceptable in R 
Examples are:
Main: The title for the plot(Displaced at the top)
Sub: The subtitle for the plot(Displaced at the bottom)
Xlim: The X-axis scale; uses the format c(min,max);automatically determined by default
ylim: The Y-axis; uses the format c(min,max):automatically determined by default
xlab: The X-axis title
ylab: the Y-axis title

col: Determines the colours used for points and lines; accepts the charcter strings of colour names(red,green,e.t.c.)
pch: The type of point to use (circle, square, triangle, e.t.c.); accepts values 0 to 25 for symbols and 32 to 255 for characters
cex: The amount to scale the size of points; accepts a numeric value; default is 1
lty: Defines the line type; accepts various characters strings(solids and dashed,dotted,e.t.c.)
Lwd: Defines the line width; accepts a positive number; default is 1

```{r}
x<-rnorm(200)
plot(x)
```

PLOTTING a scatter graph with a sequence of numbers between -5 and 5, incrementing by 0.1, choose the mean as 2 and standard devaition 0.8

```{r}
x<- seq(-5,5, by =0.1)
y<- dnorm(x, mean=2, sd=0.8)
y
plot(x,y)
```
THE FOLLOWING CODE WILL CREATE A SCATTER PLOT FOR THE RELATIONSHIP BETWEEN WT(WEIGHT) AND MPG(MILLESPERGALLON) USING MTCARS DATASET
```{r}
data(mtcars)
plot(x=mtcars$wt, y=mtcars$mpg,
    main = "scatterplot", xlab ="car weight", ylab= "miles per gallon")
mtcars
```
```{r}
plot(x=mtcars$wt,
     y=mtcars$mpg,
    main = "scatterplot", 
    xlab ="car weight",
    ylab= "miles per gallon",
    xlim =c(2.5,5),
    ylim= c(10,30),
    pch=25,
    col ="blue")
```
```{r}
iris
```

```{r}
iris
names(iris)
iris$Species
```
```{r}
iris
plot(data =iris,
     main = "Iris Data",
     Sepal.Length ~ Petal.Length,
     xlab = "petal length (cm)",
     ylab = "Sepal length (cm)",
     pch= c(16, 17, 18) [as.numeric(Species)],
     col = c("magenta", "purple", "blue") [as.numeric(Species)])
```
```{r}
x<- rnorm(100)
y<- 3 * x + rnorm(100)
z<- sqrt(1+y^2) + rnorm(100)
plot(x,y,
     main = "conditional scatter plot",
     pch = ifelse(x* y > 1,10,2),
     xlim= range(x),
     ylim= range(y,z),
     xlab = "x",
     ylab = 'Value',
     col = ifelse(y>= mean(y),'blue','magenta'))
     
     
```

simple bar chart
```{r}
mtcars
Counts <- table(mtcars$gear)
Counts
```
```{r}
barplot(Counts, main = "car distribution by gear",
        xlab="Number of Gears")
```

```{r}
barplot(Counts, main = "car distribution by gear",
        ylab="Number of Gears", horiz = TRUE)
```
```{r}
mtcars
Counts <- table(mtcars$vs,mtcars$gear)
Counts
```
```{r}
barplot(Counts, main = "car distribution by gear and vs",
       beside = TRUE,
         ylab="Number of Gears", 
         xlab = "frequency",
        col= c("lightblue", "maroon"),
        legend = rownames(Counts))
```
```{r}
colors <- c("green", "Orange", "brown")
months <- c("mar", "apr", "may","jul","sept")
Regions <- c("East","West","North")

Values <- matrix(c(2,9,3,11,9,4,8,7,3,12,5,2,8,10,11), nrow=3,
                 ncol = 5, byrow = TRUE)
barplot(Values, main= "total revenue", xlab= "months",ylab="revenue", col=colors,names.arg = months,beside=TRUE,las=2,border = "purple", font.axis= 4,col.axis = "magenta", cex.axis = 0.9,col.lab="blue", font.lab=3)
        legend("topleft",Regions, cex =0.9,fill =colors, )

        Values
```
SIMPLE PIE CHART
```{r}
Percentage <- c (10,15,38,28,10)
labels <- c("0-8 yrs 10%","9-11yrs 15%","12yrs 38%", "13-15yrs 28%","16 and above yrs 10%")
pie(Percentage,labels = labels, main = "Education percentage", col=c("Blue","magenta", "turquoise","red","purple"))
```

```{r}
X<-c(21,62,10,53)

labels <- c("LONDON","NEW YORK","SIGNAPORE","MUMBAI")
PIPEPERCENT <- round(100*X/sum(X),1)
PIPEPERCENT

pie(X, labels = PIPEPERCENT, main= "City pie chart",col= rainbow(length(X)))

legend("bottomright", c("LONDON","NEW YORK","SINGAPORE", "MUMBAI"),
       cex=0.7, fill =rainbow(length(X)))

```
```{r}
library(plotrix)
```
```{r}
library(ggplot2)
```

```{r}
X<-c(21,62,10,53)

labels <- c("LONDON","NEW YORK","SIGNAPORE","MUMBAI")
PIPEPERCENT <- round(100*X/sum(X),1)
PIPEPERCENT

pie3D(X, labels = PIPEPERCENT, main= "City pie chart",col= rainbow(length(X)))

legend("topright", c("LONDON","NEW YORK","SINGAPORE", "MUMBAI"),
       cex=0.7, fill =rainbow(length(X)))
```

LINE CHART
```{r}
V<- c(9,12,17,23,21,30,35,31,36)
X<- c(1,2,3,4,5,6,7,8,9)
plot(X,V, type= "o", col="red",xlab ="month", ylab= "Growth",main ="Growth Rate")
```

HISTOGRAM
```{r}
random = rnorm(100)
hist(random)
```

```{r}
AirPassengers
hist(AirPassengers, main = "histogram for air passengers",
     xlab = "passengers",
     border = "blue",
     col= "yellow",
     xlim= c(100,700),
     breaks=11)
```
```{r}
mpg
```


```{r}
h<-ggplot(data = mpg, aes(x = cty, y = hwy))
```


```{r}
h + geom_area(stat = "bin")
x, y, alpha, color, fill, linetype, size
```

```{r}
c + geom_area(stat = "bin")
x, y, alpha, color, fill, linetype, size
```

```{r}
economics
names(economics)
```
```{r}
h<-ggplot(data = economics, aes(x = date, y = unemploy))
h + geom_polygon(aes(date = unemploy))
```
```{r}
a<-ggplot(data = economics, aes(x = date, y = unemploy))
a + geom_path(lineend="butt", linejoin="round", linemitre=1)
a + geom_ribbon(aes(ymin=unemploy - 900, ymax=unemploy + 900))
```
```{r}
seals
b <- ggplot(seals, aes(x = long, y = lat))
b + geom_curve(aes(yend = lat + 1, xend=long+1),curvature=1)
```
```{r}
b + geom_rect(aes(xmin = long, ymin=lat, xmax= long + 1, ymax = lat + 1))

```

```{r}
c <- ggplot(mpg, aes(hwy)); c2 <- ggplot(mpg)

```

