---
title: "INTERACTIVE DASHBOARD IN R"
author: "OLUWAFIKEMI"
date: "6/3/2022"
output: html_document
---

```{r}
install.packages("flexdashboard")
```
```{r}

```


```{r}
library(highcharter)
library(ggplot2)
library(readr)
library(flexdashboard)
library(dplyr)
```


```{r}
install.packages("dplyr")
install.packages("highcharter")
install.packages("flexdashboard")
```
STEP 3, LOAD THE  DATASET INTO R
```{r}
Train<- read_csv('train.csv')
head(Train)
```
### NEXT WE USE THE THE STR() FN AND CHECK WHAT DATA TYPE THE COLUMNS ARE
```{r}
str(Train) #check column data type are colums
```
## next tell r to cast the column called pclass as a factor
```{r}
Train$Pclass = factor(Train$Pclass,labels = c("1st", "2nd","3rd"))
str(Train) #cHECK DATA TYPE ONCE AGAIN
```
## 4, names used to know the title of each columns
```{r}
names(Train)
Train
```
## 4, TELL R TO CAST THE COLUMN CALLED EMBARKED AS A FACTOR ALSO NA
```{r}
Train$Embarked <- factor(Train$Embarked, labels = c("Cherbourg","Queenstown","Southampton"))
## ib-ibadan, lag-lagos, s- southampton
```
lets check the number of rows and columns
```{r}
nrow(Train)
ncol(Train)
```
## let me check if we have missing data, like how many missing info
```{r}
colSums(is.na(Train))
```
## passenger of the titanic
## age
```{r}
tmp_males = Train %>% filter(Sex == "male",!is.na(Age))
unique(tmp_males$Sex)
nrow(tmp_males)
```
```{r}
tmp_fmales = Train %>% filter(Sex == "female",!is.na(Age))
unique(tmp_fmales$Sex)
nrow(tmp_fmales)
```

```{r}
install.packages("hrbrthemes")
install.packages("plotly")
install.packages("plyr")
```

```{r}
x<- letters[1:3]  #create an example vector
#print example vector
x
#example data is a character vector containing the three letters a,b,c.
```
```{r}
#rep fn is using times arguement
rep(x, times =3)
```
```{r}
rep(x, each =3) #apply rep with each arguement
#each vector (or list) element is repeated several times before the next element is repeated
```
```{r}
rep(x, len =9)
#apply rep with len arguement
```
```{r}
tmp_male <- Train %>% filter(Sex == "male", !is.na(Age)) %>%
  select(Age) %>% .[[1]]
tmp_male

b<- hist(tmp_male,plot = TRUE)
```
```{r}
tmp_female <- Train %>% filter(Sex == "female", !is.na(Age)) %>%
  select(Age) %>% .[[1]]
tmp_female

a<- hist(tmp_female, breaks= b$breaks,plot = TRUE)
```
```{r}
df <- data.frame(Age=c(a$mids,b$mids),Density=c(a$density,b$density),Sex= c(rep("female",length(a$mids)),rep("male",length(b$mids))))
df
```
```{r}
tmp_male <- Train %>% filter(Sex == "male", !is.na(Age)) %>%
  select(Age) %>% .[[1]]
tmp_male

b<- hist(tmp_male, 20,plot = TRUE)
```
```{r}
tmp_male <- Train %>% filter(Sex == "male", !is.na(Age)) %>%
  select(Age) %>% .[[1]]
tmp_male

b<- hist(tmp_male, 20,plot = FALSE)

#filter out female
tmp_female <- Train %>% filter(Sex == "female", !is.na(Age)) %>%
  select(Age) %>% .[[1]]
a<- hist(tmp_female, breaks = b$breaks,plot = FALSE)


df <- data.frame(Age=c(a$mids,b$mids),Density=c(a$density,b$density),Sex= c(rep("female",length(a$mids)),rep("male",length(b$mids))))


highchart() %>%
  hc_add_series(name = "female",
    select(filter(df, Sex == "female"), Density)[[1]], type = "column",
    color ='#008000', showInLegend = FALSE) %>%
  hc_add_series(name = "male",
        select(filter(df, Sex =="male"), Density) [[1]], type = "column",
        color = '#FFFF00', showInLegend = FALSE) %>%
 hc_add_series(name = "male",
            select(filter(df, Sex == "male"), Density) [[1]], type = "spline", color = "#FFFF00") %>% 
 hc_add_series(name = "female",
   select(filter(df, Sex == "female"), Density)[[1]], type = "spline",
  color ='#008000') %>%
  
  hc_tooltip(pointFormat = "<spanstyle =\"color:{series.color}\">{series.name}</span>:{point.y:.3f}<br/>", shared = FALSE) %>% 
  hc_yAxis(title = list(text = 'Density')) %>%
  hc_xAxis(title = list(text = 'Age'))
  
```

```{r}
head(Train)
```
```{r}
## ROW 3 {DATA - HEIGHT =400}
## PORT OF EMBARKATION
Tmp<- Train %>% filter(!(Embarked =="")) %>% group_by(Embarked) %>% tally() %>% mutate(Percent = n/sum(n))
Tmp
```
```{r}
Tmp <- arrange(Tmp,desc(Percent))
Tmp
```
```{r}
Tmp<- Train %>% filter(!(Embarked =="")) %>% group_by(Embarked) %>% tally() %>% mutate(Percent = n/sum(n))
Tmp$colors <- c("#d35400","#2980b9","#2ecc71")
Tmp<- arrange(Tmp, desc(Percent))
highchart() %>%
  hc_xAxis(categories = c("Southhampton", "Cherbourg", "Queenstown")) %>%
  hc_yAxis(title = list (text = 'Percentage')) %>%
  hc_add_series(Tmp, "bar", hcaes(x=Embarked, y= Percent, color =colors)) %>%
  hc_tooltip(pointFormat = "{point.y:.2f}</br>")%>%
  hc_legend(enabled = TRUE)
```

```{r}
Tmp<- Train %>% group_by(Pclass) %>% summarize(Survived = mean(Survived))
Tmp$colors <- c("#d35400", "#2980b9", "#2ecc71")
hchart(Tmp, "column",hcaes(x = Pclass, y= Survived, color = colors)) %>%
  hc_tooltip(pointFormat = "{point.y:2f}</br>", shared = FALSE)
```

### SEX (GROUPING BY SEX)
```{r}
TMP<- Train %>% group_by(Sex) %>% tally() %>% mutate(pct =n/sum(n))
TMP
```
```{r}
TMP<- Train %>% group_by(Sex) %>% tally() %>% mutate(pct =n/sum(n))
TMP$colors <- c("#d35400", "#2980b9")

hchart (TMP, "pie", hcaes(x= Sex, y= pct, color = colors)) %>% hc_tooltip(pointFormat = "{point.y:.2f}</br>", shared = FALSE)
```

KERNEL FINISHED
```{r}
rate <- 39
gauge(rate, min =0, max =100, symbol = '%', gaugeSectors(success = c(80,100), warning= c(40,79),danger=c(0,39)))
```
```{r}
gauge(39, min =0, max =75, symbol = '%', gaugeSectors(success = c(80,100), warning= c(40,79),danger=c(0,39)))
```



