---
title: "LOOPS ANALYSIS"
author: "OLUWAFIKEMI"
date: "5/18/2022"
output: html_document
---
*LAPY(): Loop over a list of vector and evaluate a function on each element
*SAPLY(): same as laply but try to simplify the result
*APPLY(): Apply a function over the margins of an array,matrics or data frames
*TAPLY(): Apply a function over subsets of a vector
*MAPLY(): Multivariate version of lapply
\

            lapply()
The lapply() function does the following simple series of operations:
1. It loops over a list, iterating (REPEAT OVER AND OVER THE SAME PROCESS) over each element in that list
2. It applies a function to each element of the list(a function that you specify)
3. and returns a list (the 1 is for "list").

This function takes the argument:
(1) a list X;
(2) A function (or the name of a function)  FUN;
(3) Other arguments via its...arguments
If X is not a list, it will be coerced to a list using as.list()

```{r}
mat= cbind (x=c(3,0,3,3), y=c(3,0,0,0), z=c(3,0,0,3),r= c(1,1,0,0),d=c(1,1,1,0), v=c(1,1,1,0))
mat
matr=rbind(c(3,0,3,3), c(3,0,0,0), c(3,0,0,3), c(1,1,0,0),c(1,1,1,0), c(1,1,1,0))
matr
class(mat)
```

```{r}
make <- apply(mat,1,sum)
makes <- apply(mat,2,sum)
make
makes
```
```{r}
mak <- apply(mat,1,min)
ma <- apply(mat,2,min)
mak
ma
```
```{r}
made <- apply(mat,1,max)
may <- apply(mat,2,max)
made
may
```
```{r}
a=1:5
a
b=rnorm(10)
```
```{r}
rnorm(10,5)
```
```{r}
x<-list(a=1:5,b=rnorm(10))
x
```

```{r}
lapply(x,mean)
```
```{r}
input=c(1,2,3)
lapply(input,function(x) x^2)
```
```{r}
input=c(1,2,3)
lapply(input,function(x) x+2)
```
```{r}
x<-list(a=1:4, b=rnorm(10),c=rnorm(20,1), d=rnorm(100,5))
x
lapply(x,sum)
lapply(x, mean)
lapply(x,median)
```
```{r}
#x<-list(a=1:4, b=rnorm(10),c=rnorm(20,1), d=rnorm(100,5))
#x
lapply(x,function(y) y^2)
lapply(x,function(y) y/5)
lapply(x,function(y) y*10)
```
you can use lapply() to evaluate a function multiple times each with a different argument.Below is an example where i call the runif() function (to generate uniformly distributed random variable) four times, each time generating a different  number of random numbers.
```{r}
runif(1)
runif(2)
runif(5)
runif(20,0.3,0.6)
```

```{r}
x<-1:4
x
lapply(x,runif)  #run random number sequentially
```
```{r}
x=1:4
lapply(x,runif,min=0, max= 10) #passing multiple  func min and max
#default random number runif is 0 and 1 but now it is between 0 to  10
```
```{r}
x<- list(a=matrix(1:4, 2,2), b=matrix(1:6,3,2))
x
```
```{r}
lapply(x, function(elt) { elt[,2]})
lapply(x, function(elt) { elt[2,]})
lapply(x, function(elt) { elt[1,2]})
#elt fn helps to pick a particular row or column from a lappy
```
#maply it can apply a func on an unlimited set of vectors/list, it is like a version of lapply that can handle multiple vectors as argument
```{r}
XS =0:5
YS= c(2,2,2,3,3,3)
results<-mapply(function(x,y) sum(x,y),XS,YS)
results
```
```{r}
x<- list(a=1:4,b =rnorm(10),c=rnorm(20,1),d=rnorm(100,5))
results <-mapply(sum,x)
results
```
```{r}

#Create Vector quantity
quantity <- 31
#Create multiple condition statement
if (quantity <= 20){
  print("Not Enough for Today")
} else if (quantity > 20 & quantity <= 30){
  print("Average Day")
} else{
  print ("What a Great Day!")
}
```
```{r}

category <- 'A'
Price <- 10
if (category == 'A') {
  cat('A Vat rate of 8% is applied.', 'The total price is', Price * 1.08)
} else if (category == 'B') { cat('A Vat rate of 10% is applied.', 'The total price is', Price * 1.10)
} else if (category == 'C') { cat('A Vat rate of 15% is applied.', 'The total price is', Price * 1.15) } else { cat('A Vat rate of 20% is applied.', 'The total price is', Price * 1.20) 
}

```





