---
title: "MANIPULATING DATA USING GGPLOT2"
author: "OLUWAFIKEMI"
date: "6/18/2022"
output: html_document
---
---
title: "MANIPULATING GGPLOT USING HR DATASET"
author: "OLUWAFIKEMI"
date: "6/10/2022"
output: html_document
---

```{r}
library(readr)
HR_Dataset <- read_csv("HR Dataset.csv")
head(HR_Dataset)
```

```{r}
library(highcharter)
library(ggplot2)
library(readr)
library(flexdashboard)
library(dplyr)
library(plotly)
```
```{r}
library(plotly)
```


```{r}
DATA = HR_Dataset
DATA
```
```{r}
unique(DATA$STATUS)
```

# DATA WRANGLING - CHECK IF THERE ARE MISSING VALUES OR INFO
```{r}
sum(is.null(DATA))
```
#count no if rows we have
```{r}
nrow(DATA)
```
```{r}
summary(DATA)
```
```{r}
ncol(DATA)
```
#Earlier we had indicated that we had both active records at end of year and terminates during the year for each of 10years going from 2006 to 2015. To have a population to model from(to differentiate ACTIVES FROM TERMINATES) We have to include both status types
Its useful then to get a baseline of what percent/ porportion the terminates are  of the entire population.It also answers our first question.Let's look at the next.

```{r}
StatusCount = as.data.frame.matrix(DATA %>%
  group_by(STATUS_YEAR)%>%
  select(STATUS)%>% table())

#StatusCount
StatusCount$TOTAL <- StatusCount$ACTIVE + StatusCount$TERMINATED
StatusCount$PercentTerminated <- StatusCount$TERMINATED/(StatusCount$TOTAL)*100
StatusCount$PercentACTIVE <- StatusCount$ACTIVE/(StatusCount$TOTAL) *100
StatusCount
mean(StatusCount$PercentTerminated)
mean(StatusCount$PercentACTIVE)
```
```{r}
StatusCount = as.data.frame.matrix(DATA) %>%
  group_by(STATUS_YEAR)%>%
  select(STATUS)%>% table()
```
```{r}
#library(ggplot2
ggplot()+ geom_bar (aes( y=  ..count.. , x =
as.factor(BUSINESS_UNIT),fill=(STATUS)),
data = DATA,position = position_stack())

```
```{r}
ggplot(DATA,aes(y=..count..,x= as.factor(BUSINESS_UNIT),fill=(STATUS)))+geom_bar()
```
```{r}
h<-ggplot()+ geom_bar (aes( y=  ..count.. , x =
as.factor(BUSINESS_UNIT),fill=(STATUS)),
data = DATA,position = position_dodge2())
ggplotly(h,tooltip = "y")
```
```{r}
Job_Termination
```
# Job Termination findings according to termination type and status year
```{r}
Job_Termination = as.data.frame(DATA %>% filter(STATUS=="TERMINATED")) #== MEANS FILTER
#Job_Termination
h= ggplot()+ geom_bar(aes(y=..count..,x= as.factor(STATUS_YEAR),fill=as.factor(termtype_desc)),data = Job_Termination,position = position_stack())

ggplotly(h,tooltip = "y")
h
```
#just terminate by status year and termination reas
```{r}
h= ggplot()+ geom_bar(aes(y=..count..,x= as.factor(STATUS_YEAR),fill=as.factor(termreason_desc)),data = Job_Termination,position = position_stack())
h
```
#looking at job termination report by termination reason and department
```{r}
h= ggplot()+ geom_bar(aes(y=..count..,x= as.factor(department_name),fill=as.factor(termreason_desc)),data = Job_Termination,position = position_stack())+ theme(axis.text = element_text(angle = 90,hjust=1,vjust = 0.3))
ggplotly(h,tooltip = "y")
```
#generallly most termination seem to be voluntary year by year except in the most recent year where are some involuntary termination
```{r}
?hrbrthemes
```

```{r}
Data_Analysts= c(rep("Yemi",3), rep("Kemi",3),rep("Hassan",3),rep("Tony",3),rep("Abdul",3),rep("Bisola",3)) #rep stands for repeat times
Data_Analysts
```
```{r}
skills = rep(c("excel","R","power Bi"),6) #rep stands for repeat items
skills
```
```{r}
scores = abs(rnorm(18,0,20))
scores
```
```{r}
data <- data.frame(Data_Analysts,skills,scores) #grouped to table format
data
```
```{r}
p=ggplot(data,aes(fill=skills,y=scores,x=Data_Analysts))+ geom_bar(position="Stack", stat = "identity")
## we have stack and dodge position
ggplotly(p, tooltip="y")
```

