---
title: "MULTIPLE LINEAR REGRESSION"
author: "OLUWAFIKEMI"
date: "6/20/2022"
output: html_document
---
#Supervised learning model with linear regression
#_our_baby_steps_
collecting the data in R
Checking for linearity
Applying the multiple linear regression model
making a prediction
----linear regression is used to make prediction

STEP 1: COLLECT THE DATA

So let"s start with a simple example where the goal is to predict the stock index price(the depedent variable)of a fictious economy based on two independent / input variables:

interest rate- independent variable
unemployment rate - independent variable
depended variable - stock price is what we wish to predict

#lets create a small data to use
```{r}
year<-c(2017,2017,2017,2017,2017,2017,2017,2017,2017,2017,2017,2017,2016,2016,2016,2016,2016,2016,2016,2016,2016,2016,2016,2016)
length(year)
```


```{r}
month<-c(12,11,10,9,8,7,6,5,4,3,2,1,12,11,10,9,8,7,6,5,4,3,2,1)
month
length(month)
```
```{r}
interest_rate<-c(2.75,2.5,2.5,2.5,2.5,2.5,2.5,2.25,2.25,2.25,2,2,2,1.75,1.75,1.75,1.75,1.75,1.75,1.75,1.75,1.75,1.75,1.75)
length(interest_rate)
```

```{r}
unemployment_rate<-c(5.3,5.3,5.3,5.3,5.4,5.6,5.5,5.5,5.5,5.6,5.7,5.9,6,5.9,5.8,6.1,6.2,6.1,6.1,6.1,5.9,6.2,6.2,6.1)
length(unemployment_rate)
```
```{r}
stock_index_price <-c(1464,1394,1357,1293,1256,1254,1234,1195,1159,1167,1130,1075,1047,965,943,958,971,949,884,866,876,822,704,719)
length(stock_index_price)
```

```{r}
interest_rate
is.data.frame(interest_rate)
nrow(interest_rate)
```

```{r}
getwd()
```
THIS IS HOW YOU PUT IT IN A DATA FRAME
```{r}
df = do.call(rbind,Map(data.frame, years = year, months = month, unemployement_rates=unemployment_rate, stock_index_prices=stock_index_price,interest_rates=interest_rate))
df
```
```{r}
class(df$years)
class(df$months)
class(df)
typeof(df)
str(df)
```
```{r}
dim(df) #how many rows and columns, IN R, rows is observation and columns is variables
```
```{r}
df[c('unemployement_rates','years')]
```
```{r}
nrow(df['stock_index_prices'])
```
```{r}
df["years"]
```
STEP 3: OKAY NA, LETS CHECK FOR LINARITY
Before you apply linear regression models, you will need to verify that several assumptions are met.Most notably, you'll need to make sure that a linear relationship exists between the dependent variable and the independent variable/s

A quick way to check for linearity is by using scatter plots.
For our example, we will check that a linear relationship exists between:

#The stock_index_price(dependent variable) and the interest_rate( independent variable); and 

the  stock_index_price (dependent variable) and unemployment_rate(independent variable)

```{r}
max(df['stock_index_prices']) #check the max price
min(df['stock_index_prices'])
```
lets use plot interest vs stock price
```{r}
plot(x=interest_rate, y= stock_index_price)
#there is +ve linearity
```

```{r}
plot(x=unemployment_rate, y= stock_index_price)
#there is +ve linearity
```
#export the data frame from R to excel
```{r}
write.csv(df,"liner reg.csv")
```

```{r}
cor(stock_index_price,interest_rate)
```
```{r}
cor(stock_index_price,unemployment_rate)
#notes, omg! it seems like we are fuck,meaning there's no no
#fucking relationship between stock price and unemploymment
#practically, as stock price goes up, unemployment_rate goes down
```
Notes you will notice that indeed a linear relationship exists between the stock_interest_rate, specifically, when interest rate go up, the stock index price also goes up:


As you can see a linear relationship also exists between the stock_index_price and the unemployment rate - when the unemployment rate go up, the stock index price goes down (here we still have a linear relationship, but with a negative slope)

STEP 4 Apply the multiple linear regression in R
```{r}
# model = lm(dependent variable ~ first inependent variable + second independent variable+...)
#summary(model)

```
This is where we need to start training our linear regression model- learning the data
```{r}
model=lm(stock_index_price ~ interest_rate + unemployment_rate)
summary(model)
model
```

You can use the coefficient in the summary in order to build the multiple linear regression equation as follows:

stock_index_price = (intercept) + (interest_rate coef)*X1
(unemployment_rate coef)*X2
X1 means the unknown interesr rate
X2 means the unknown unmployment rate

And once you plug the numbers from the summary
stock_index_price = (1798.4) + (345.5)* X1 + (-250.1)*x2

STEP 5: MAKE A PREDICTION
Now lets make a prediction based on the equation above.
For example, imagine that you want to predict the stock index price after you collected the following data

interest rate = 1.5(i.e., X1 = 1.5)
Uenmployment rate = 5.8(i.e., X2 = 5.8)

And if you plug that data into the regression equation you will get:
stock_index_price = (1798.4) + (345.5)* (1.5) + (-250.1)* (5.8) = 866.07

y= c+ m1X1 +m2X2
Y= m1X1 +m2X2 + m3X3 + .....+C

```{r}
stock_index_price =(1798.4) + (345.5)*(1.5) + (-250.1)*(5.8)
stock_index_price
```
The predicted value for the stoc_index_price is therefore 866.07
