---
title: "REGRESSION TREE"
author: "OLUWAFIKEMI"
date: "6/22/2022"
output: pdf_document
---
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(plotly)
library(rattle)
library(caret)
```


```{r}
install.packages("rattle")
install.packages("caret")
```
```{r}
getwd()
```
```{r}
library(readr)
TRUCKLOAD_ANALYSIS <- read_csv("C:\\Users\\acer\\Documents\\R\\TRUCKLOAD ANALYSIS.csv")
head(TRUCKLOAD_ANALYSIS)
```
```{r}
kemidata= TRUCKLOAD_ANALYSIS
kemidata
```
you can also import data by typing this down here but mak sure the data you want to import is in the working direction you are working with
```{r}
data= read.csv("TRUCKLOAD_ANALYSIS")
data
```

```{r}
class(kemidata)
```
```{r}
str(kemidata)
```
```{r}
names(kemidata)
```

```{r}
unique(kemidata$Month)
unique(kemidata$Year)
unique(kemidata$`Truckout date`)
```
```{r}
truck_data_picked <- kemidata %>% select(c("Truckout distribution", "Quantity loaded","Destination state","Month", "Year" ))
truck_data_picked
```
```{r}
class(truck_data_picked)
```
```{r}
str(truck_data_picked)
```
```{r}
summary(truck_data_picked)
```
```{r}
training_data_jan_march <- mytrain<- subset(truck_data_picked, Month <= "March")
TESTDATA_APRIL = subset(truck_data_picked, Month == "April")
TESTDATA_APRIL
#you can also use the syntax select,filter or subset to filter
```
YOU CAN EXPORT THE DATA TO EXCEL TO CHECK THE FILTERING

```{r}

#NOW EXPORT THEM TO CSV
write.csv(mytrain,'TRUCKTRAIN DATA.CSV')
write.csv(TESTDATA_APRIL,'TRUCKTEST DATA.CSV')
```

```{r}
mytrain
training_data_jan_march
```
```{r}
library(rpart)
library(rpart.plot)
```

```{r}
TRUCKED_FINISHED = rpart(training_data_jan_march$`Truckout distribution` ~ training_data_jan_march$`Quantity loaded`+ training_data_jan_march$Month, data= (training_data_jan_march))
TRUCKED_FINISHED
```
```{r}
rpart.plot(TRUCKED_FINISHED)
```
```{r}
PREDICTING <- predict(TRUCKED_FINISHED,TESTDATA_APRIL)
PREDICTING
summary(PREDICTING)
```


```
```{r}
install.packages("rattle")
```


```{r}
install.packages("rattle", repos = "rattle.togaware.com",type = "source")
```

```{r}
install.packages("https://cran.r-project.org/bin/window/contrib/3.3/RGtk2_2.20.31.zip",repose = NULL)
```



```{r}
Sys.which("make")
```

```{r}
#remove.packages("rattle") to remove packages
```

```{r}
install.packages("https://access.togaware.com/RGtk2_2.20.36.2.zip", repos=NULL)
```

```{r}
rattle()
```

